Generalization Errors of Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-09-09 arxiv/proof_main.tex
pf overview

Before the end of this section, we describe briefly the main ideas in the proofs of Lemmas \ref{lem_cov_shift} and \ref{lem_cov_derivative}.

\medskip
\noindent\textbf{An overview of the proofs of Lemmas \ref{lem_cov_shift} and \ref{lem_cov_derivative}.}
%We first describe the proof of Theorem \ref{lem_cov_shift_informal}.
%We briefly describe the key ideas for proving Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative}.
We use the standard Stieltjes transform method (or the resolvent method) in random matrix theory \cite{bai2009spectral,tao2012topics,erdos2017dynamical}. Both Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative} require a study of the matrix inverse $(X_1^\top X_1+X_2^\top X_2)^{-1}$ for $X_1= Z_1\Sigma_1^{1/2}$ and $X_2= Z_2\Sigma_2^{1/2}$. Suppose $M=\Sig_1^{1/2} \Sig_2^{-1/2}$ has a singular value decomposition
\be\label{eigen2}
M= U\Lambda V^\top, \quad \Lambda=\text{diag}( \lambda_1, \ldots, \lambda_p).
\ee
Then using equation \eqref{eigen2}, we can write
\be\label{eigen2extra}(X_1^\top X_1+X_2^\top X_2)^{-1}= \Sigma_2^{-1/2}V\left(   \Lambda U^\top Z_1^\top Z_1 U\Lambda  + V^\top Z_2^\top Z_2V\right)^{-1}V^\top\Sigma_2^{-1/2}.\ee
We divide the proof into the following parts.

@@

\medskip
\noindent\textbf{Defining the asymptotic limits of the resolvents.} 
%\label{sec aymp_limit_G}
%\subsection{Resolvents and limiting law}
We now describe the asymptotic limit of $G(z)$ as $n\to \infty$. First we define the deterministic limits of $(m_1(z), m_{2}(z))$ by $\left(\frac{\rho_1+\rho_2}{\rho_1}a_{1}(z),\frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)\right)$, where $(a_1(z), a_2(z))$ is
%\HZ{This $M_{1}, M_{2}$ notation is awkward; consider changing it to ${M}_1(z), {M}_2(z)$ (or  better, say $a_1(z), a_2(z)$? if they correspond to $a_1, a_2$ when $z=0$)}, 
the unique solution to the following system of equations
\begin{equation}\label{selfomega}
\begin{split}
& \frac{r_1}{a_{1}(z)} = \frac{1}{n}\sum_{i=1}^p \frac{\lambda_i^2}{  z+\lambda_i^2 a_{1}(z) +a_{2} (z) } - 1 ,\ \  \ \frac{r_2}{a_{2}(z)} = \frac{1}{n}\sum_{i=1}^p \frac{1 }{  z+\lambda_i^2 a_{1}(z) +  a_{2}(z)  }- 1 ,
\end{split}
\ee
satisfying that $\im a_{1}(z)>0$ and $\im a_{2}(z)>0$ for $z\in \C_+$ with $\im z$. 
%Here, for simplicity of notations, we introduced the following ratios 
%\be\label{ratios}
% \gamma_n :=\frac{p}{n}=\frac{1}{\rho_1+\rho_2},\quad r_1 :=\frac{n_1}{n}=\frac{\rho_1}{\rho_1+\rho_2},\quad r_2 :=\frac{n_2}{n}=\frac{\rho_2}{\rho_1+\rho_2}.
%\ee
We define the matrix limit of $G(z)$ as
\be \label{defn_piw}
\Pi(z) := \begin{pmatrix} -(z+a_{1}(z)\Lambda^2  +  a_{2}(z))^{-1} & 0 & 0 \\ 0 &  \frac{\rho_1+\rho_2}{\rho_1} a_{1}(z)\id_{n_1} & 0 \\ 0 & 0 & \frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)\id_{n_2}  \end{pmatrix}.\ee
In particular, the matrix limit of $\cal G(z)$ is given by $-(z+a_{1}(z)\Lambda^2 + a_{2}(z))^{-1}$.

\medskip
\noindent{\bf Deriving the bias and variance asymptotics.} {\cor Notice that using equation \eqref{mainG}, we can write  equation \eqref{eigen2extra} as
\be\label{rewrite X as R} (X_1^\top X_1+X_2^\top X_2)^{-1}=n^{-1}\Sigma_2^{-1/2}V\cal G(0)V^\top\Sigma_2^{-1/2}.\ee
For the bias asymptotic limit, we show in \todo{where?} that
$$\bignorm{\Sigma_2^{1/2} (X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} w}^2= w^\top \Sigma_2^{-1/2}R'(0)\Sigma_2^{-1/2} w.$$

In the above definition, we have taken the argument of $\cal G$ to be a general complex number, because we will need to use $\cal G'(0)$ in the proof of Lemma \ref{lem_cov_derivative}, which requires a good estimate of $\cal G(z)$ for $z$ around the origin.

If $z=0$, then equation \eqref{selfomega} can be reduced to 
\begin{equation}\label{selfomega0}
\begin{split}
&a_1 + a_2= 1-\gamma_n,\quad    a_1 +\frac{\gamma_n}p\sum_{i=1}^p \frac{\lambda_i^2 a_1}{ \lambda_i^2 a_1+ (1-\gamma_n - a_1) }=r_1 .
\end{split}
\ee
where $a_1:=-r_1M_{1}(0)$ and $a_2:=-r_2M_{2}(0)$.

For the variance asymptotic limit, we study the resolvent
	\[ R(z):= \bigbrace{\Sigma_2^{-1/2}( X_1^{\top}X_1 + X_2^{\top}X_2)\Sigma_2^{-1/2} - z \id}^{-1}, \text{ for any } z\in \C \text{ around } z=0. \]
Using the techniques from \citet{Anisotropic} and \citet{yang2019spiked}, we find the asymptotic limit of $R(z)$ for any $z$ as $p$ goes to infinity, denoted by $R_\infty(z)$, with an almost optimal convergence rate of $p$.
In particular, when $z=0$, the asymptotic limit of equation \eqref{lem_cov_shift_eq} is given by
	\[ \tr[\Sigma_2^{-1/2} \Sigma \Sigma_2^{-1/2}R_\infty(0)]. \]

For the bias asymptotic limit, we show in \todo{where?} that
$$\bignorm{\Sigma_2^{1/2} (X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} w}^2= w^\top \Sigma_2^{-1/2}R'(0)\Sigma_2^{-1/2} w.$$
Hence its limit can be calculated through $R_\infty'(z)$, which gives the expression in \eqref{lem_cov_derv_eq}.
We leave the full proof of Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative} to Appendix \ref{sec_maintools}.
Combining the above two results, we provide the proof of Theorem \ref{thm_main_informal} in Section \ref{app_proof_main_thm}.
}

\medskip
\noindent{\bf Defining the resolvent minor.} \ 
%\HZ{Divide into several parts like sec 7 to improve readability.}
Then we define the resolvent minors, which are defined by removing certain rows and columns of the matrix $H$.

\begin{definition}[Resolvent minors]\label{defn_Minor}
 For any $ (p+n)\times (p+n)$ matrix $\cal A$ and $\fa \in \mathcal I$, we define the minor $\cal A^{(\mathfrak c)}:=(\cal A_{\fa\fb}:\fa,\fb \in \mathcal I\setminus \{\mathfrak a\})$ as the $ (p+n-1)\times (p+n-1)$ matrix obtained by removing the $\mathfrak c$-th row and column in $\cal A$. Note that we keep the names of indices when defining $\cal A^{(\mathfrak c)}$, i.e. $(\cal A^{(\mathfrak c)})_{\fa\fb}= \cal A_{\fa\fb}$ for $\fa,\fb \ne \mathfrak c$. Correspondingly, we define the resolvent minor as %(recall equation \eqref{green2})
\begin{align*}
G^{(\mathfrak c)}:&=\left[\left(H - \left( {\begin{array}{*{20}c}
   { zI_{p}} & 0  \\
   0 & { I_{n}}  \\
\end{array}} \right)\right)^{(\mathfrak c)}\right]^{-1} ,
%= \left( {\begin{array}{*{20}c}
%   { \mathcal G^{(\mathbb T)}} & \mathcal G^{(\mathbb T)} W^{(\mathbb T)}  \\
%   {\left(W^{(\mathbb T)}\right)^\top\mathcal G^{(\mathbb T)}} & { \mathcal G_R^{(\mathbb T)} }  \\
%\end{array}} \right)  ,
\end{align*}
and define the partial traces $m^{(\mathfrak c)}$ and $m^{(\mathfrak c)}_k$, $k=0,1,2,$ by replacing $G$ with $G^{(\mathfrak c)}$ in equation \eqref{defm}. For convenience, we will adopt the convention that for the resolvent minor $G^{(\mathfrak c)}$ defined as above, $G^{(\mathfrak c)}_{\fa\fb} = 0$ if $\fa =\mathfrak c$ or $\fb =\mathfrak c$.
\end{definition}

Note that the resolvent minor $G^{(\mathfrak c)}$ is defined such that it is independent of the entries in the $\mathfrak c$-th row and column of $H$. One will see a crucial use of this fact in the heuristic proof below.

\medskip
\noindent{\bf Schur complements and concentration.}
Now we give a heuristic derivation of the matrix limit when the entries of $Z_1$ and $Z_2$ are i.i.d. Gaussian. Note that in this case,
%However, notice that if the entries of $Z_1\equiv Z_1^{\text{Gauss}}$ and $Z_2\equiv Z_2^{\text{Gauss}}$ are i.i.d. Gaussian, then 
by the rotational invariance of the multivariate Gaussian distribution we have
\be\label{eq in Gauss} Z_1 U\Lambda \stackrel{d}{=} Z_1 \Lambda, \quad Z_2 V \stackrel{d}{=} Z_2,\ee
where ``$\stackrel{d}{=}$" means ``equal in distribution". Hence it suffices to consider the following resolvent  
 \begin{equation} \nonumber
   G(z)= \left( {\begin{array}{*{20}c}
   { -z\id_{p} } & n^{-1/2}\Lambda Z_1^\top & n^{-1/2} Z_2^\top  \\
   {n^{-1/2} Z_1 \Lambda  } & {-\id_{n_1}} & 0 \\
   {n^{-1/2} Z_2} & 0 & {-\id_{n_2}}
   \end{array}} \right)^{-1}.
 \end{equation}
Using Schur complement formula (see \alert{equation} (\ref{resolvent2}) \HZ{check other usage of equation ref}), we have that %for $i \in \cal I_0$, $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\begin{align}
\frac{1}{{G_{ii} }}&=  - z - \frac{\lambda_i^2}{n} \sum_{\mu,\nu\in \mathcal I_1} z_{\mu i}z_{\nu i}G^{\left( i \right)}_{\mu\nu} - \frac{1}{n} \sum_{\mu,\nu\in \mathcal I_2} z_{\mu i}z_{\nu i}G^{\left( i \right)}_{\mu\nu} -2 \frac{\lambda_i}{n} \sum_{\mu\in \cal I_1,\nu\in \mathcal I_2} z_{\mu i}z_{\nu i}G^{\left( i \right)}_{\mu\nu},  \ \ \text{for }\  i \in \cal I_0 , \label{0self_Gii}\\
\frac{1}{{G_{\mu\mu} }}&=  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}\lambda_i \lambda_j z_{\mu i}z_{\mu j} G^{\left(\mu\right)}_{ij}, \quad \frac{1}{{G_{\nu\nu} }}=  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}  z_{\nu i}z_{\nu j}  G^{\left(\nu\right)}_{ij},  \ \ \text{for }\  \mu \in \cal I_1, \ \nu\in \cal I_2, \label{0self_Gmu1} 
\end{align}
where we recall the notations in equation \eqref{labelZ}. For the right-hand side of equation \eqref{0self_Gii}, $G^{(i)}$ is independent of the entries $z_{\mu i}$ and $z_{\nu i}$. Hence by the concentration inequalities in Lemma \ref{largedeviation}, we have that the  right-hand side of equation \eqref{0self_Gii} concentrates around the partial expectation over the entries $\{z_{\mu i}: \mu \in \cal I_1\cup \cal I_2\}$, i.e., with high probability,
\begin{align*}
\frac{1}{{G_{ii} }}&=  - z - \frac{\lambda_i^2}{n} \sum_{\mu \in \mathcal I_1}  G^{\left( i \right)}_{\mu\mu} - \frac{1}{n} \sum_{\mu\in \mathcal I_2} G^{\left( i \right)}_{\mu\mu} +\oo(1)= - z - \lambda_i^2 r_1 m_1^{(i)}(z)-  r_2m_2^{(i)}(z)+\oo(1),  
\end{align*}
where we used the definition of $m_1^{(i)}$ and $m_2^{(i)}$ in equation \eqref{defm} with $G$ replaced by $G^{(i)}$. Intuitively, since we have only removed only one column and one row out of the $(p+n)$ columns and rows in $H$, $m_1^{(i)}$ and $m_2^{(i)}$ should be close to the original $m_1$ and $m_2$. Hence we obtain from the above equation that
\begin{align}\label{1self_Gii}
 G_{ii}  = -\frac{1}{ z +\lambda_i^2 r_1  m_1(z) + r_2 m_2(z)+\oo(1)}.
\end{align}
Similarly, we can obtain from equation \eqref{0self_Gmu1} that for $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\be\label{1self_Gmu} G_{\mu \mu }=-\frac{1}{1+\gamma_n m_0 + \oo(1)},\quad G_{\nu\nu}=-\frac1{1+\gamma_n m+\oo(1)},\ee
with high probability.

\medskip
\noindent\textbf{Deriving the self-consistent equations.} Taking average we obtain that 
\be\label{2self_Gmu} m_1= \frac{1}{n_1}\sum_{\mu \in \cal I_1}G_{\mu\mu}=-\frac{1}{1+\gamma_n m_0 + \oo(1)},\quad m_2=\frac{1}{n_2}\sum_{\nu \in \cal I_2}G_{\nu\nu}=-\frac{1}{1+\gamma_n m + \oo(1)},
\ee
with high probability. Together with the definition of $m$ and $m_0$ in equation \eqref{defm}, the two equations in equation \eqref{2self_Gmu} give that 
\be\label{3self_Gmu}  \frac1{m_1}= -1- \frac{\gamma_n}{p} \sum_{i=1}^p \lambda_i^2 G_{ii}+ \oo(1),\quad \frac1{m_2}=-1-\frac{\gamma_n}{p} \sum_{i=1}^p G_{ii}  + \oo(1),\ee
with high probability. Plugging equation \eqref{1self_Gii} into equation \eqref{3self_Gmu}, we obtain that
\begin{align*}
& \frac1{m_1}= -1+ \frac{\gamma_n}{p} \sum_{i=1}^p \frac{\lambda_i^2 }{ z +\lambda_i^2 r_1  m_1(z) + r_2 m_2(z)+\oo(1)}+ \oo(1),\\ 
& \frac1{m_2}=-1+\frac{\gamma_n}{p} \sum_{i=1}^p \frac{1}{ z +\lambda_i^2 r_1  m_1(z) + r_2 m_2(z)+\oo(1)}  + \oo(1),
\end{align*}
with high probability, which give the approximate self-consistent equations for $(m_1,m_2)$. Compare them to the deterministic self-consistent equations in equation \eqref{selfomega}, one can observe that we should have $(m_1,m_2) =(M_{1}, M_{2})+\oo(1)$ with high probability.

\medskip
\noindent\textbf{Deriving the matrix limit.} Inserting this approximate identity into equation \eqref{1self_Gii}-equation \eqref{2self_Gmu}, we see that for  $i \in \cal I_0$, $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
$$G_{ii}=-(z +\lambda_i^2 r_1  M_{1} + r_2 M_{2}+\oo(1))^{1/2},\quad G_{\mu\mu}=M_{1}+\oo(1),\quad G_{\nu\nu}=M_{2}+\oo(1),$$
with high probability. These explain the diagonal entries of $\Pi$ in equation \eqref{defn_piw}. For the off-diagonal entries, they are close to zero due to concentration. For example, for $i\ne j\in \cal I_1$, by Schur complement formula (see (\ref{resolvent3})), we have
$$G_{ij}=-G_{ii}\left(\frac{\lambda_i}{n^{1/2}}\sum_{\mu \in \cal I_1} z_{\mu i} G^{(i)}_{\mu j} + \frac{1}{n^{1/2}}\sum_{\mu \in \cal I_2} z_{\mu i} G^{(i)}_{\mu j} \right).$$
Using Lemma \ref{largedeviation}, we can show that $n^{-1/2}\sum_{\mu \in \cal I_1} z_{\mu i} G^{(i)}_{\mu j}$ and $n^{-1/2}\sum_{\mu \in \cal I_2} z_{\mu i} G^{(i)}_{\mu j}$ are both close to zero. The other off-diagonal entries can be bounded in the same way.

The above arguments are the core of the main proof. To have a rigorous proof, we need to estimate each error carefully, and extend the Gaussian case to the more general case where the entries of $Z_1$ and $Z_2$ only satisfy certain moment assumptions. These will make the real argument rather tedious, but the methods we used are standard in the random matrix literature \cite{erdos2017dynamical,Anisotropic}. For the full proof, we refer the reader to ... .

>>>>

\subsection{Proof Overview for the Random Matrix Results}
Finally, in this subsection, we describe briefly the main ideas in the proofs of Lemmas \ref{lem_cov_shift} and \ref{lem_cov_derivative}. 
%We first describe the proof of Theorem \ref{lem_cov_shift_informal}.
%We briefly describe the key ideas for proving Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative}.
We use the standard Stieltjes transform method (or the resolvent method) in random matrix theory \cite{bai2009spectral,tao2012topics,erdos2017dynamical}. Both Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative} require a study of the matrix inverse $(X_1^\top X_1+X_2^\top X_2)^{-1}$ for $X_1= Z_1\Sigma_1^{1/2}$ and $X_2= Z_2\Sigma_2^{1/2}$. Suppose $M=\Sig_1^{1/2} \Sig_2^{-1/2}$ has a singular value decomposition
\be\label{eigen2}
M= U\Lambda V^\top, \quad \Lambda=\text{diag}( \lambda_1, \ldots, \lambda_p).
\ee
Then using equation \eqref{eigen2}, we can write
\be\label{eigen2extra}(X_1^\top X_1+X_2^\top X_2)^{-1}= \Sigma_2^{-1/2}V\left(   \Lambda U^\top Z_1^\top Z_1 U\Lambda  + V^\top Z_2^\top Z_2V\right)^{-1}V^\top\Sigma_2^{-1/2}.\ee
We divide the discussion into the following parts.

@@

\medskip
\noindent\textbf{Defining the asymptotic limits of the resolvents.} 
%\label{sec aymp_limit_G}
%\subsection{Resolvents and limiting law}
We now describe the asymptotic limit of $G(z)$ as $n\to \infty$. First we define the deterministic limits of $(m_1(z), m_{2}(z))$ by $\left(-\frac{\rho_1+\rho_2}{\rho_1}a_{1}(z),-\frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)\right)$, where $(a_1(z), a_2(z))$ is
%\HZ{This $M_{1}, M_{2}$ notation is awkward; consider changing it to ${M}_1(z), {M}_2(z)$ (or  better, say $a_1(z), a_2(z)$? if they correspond to $a_1, a_2$ when $z=0$)}, 
the unique solution to the following system of equations
\begin{equation}\label{selfomega_a}
\begin{split}
& \frac{\rho_1}{a_{1}(z)} = \frac{1}{p}\sum_{i=1}^p \frac{\lambda_i^2}{ - z+\lambda_i^2 a_{1}(z) +a_{2} (z) } + (\rho_1+\rho_2),\  \frac{\rho_2}{a_{2}(z)} = \frac{1}{p}\sum_{i=1}^p \frac{1 }{  -z+\lambda_i^2 a_{1}(z) +  a_{2}(z)  }+ (\rho_1+\rho_2) ,
\end{split}
\ee
satisfying that $\im a_{1}(z)>0$ and $\im a_{2}(z)>0$ for $z\in \C_+$ with $\im z$. 
%Here, for simplicity of notations, we introduced the following ratios 
%\be\label{ratios}
% \gamma_n :=\frac{p}{n}=\frac{1}{\rho_1+\rho_2},\quad r_1 :=\frac{n_1}{n}=\frac{\rho_1}{\rho_1+\rho_2},\quad r_2 :=\frac{n_2}{n}=\frac{\rho_2}{\rho_1+\rho_2}.
%\ee
We define the matrix limit of $G(z)$ as
\be \label{defn_piw}
\Gi(z) := \begin{pmatrix} (-z\id_p+a_{1}(z)\Lambda^2  +  a_{2}(z)\id_p)^{-1} & 0 & 0 \\ 0 & - \frac{\rho_1+\rho_2}{\rho_1} a_{1}(z)\id_{n_1} & 0 \\ 0 & 0 & -\frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)\id_{n_2}  \end{pmatrix}.\ee
In particular, the matrix limit of $\cal G(z)$ is given by $(-z\id_p+a_{1}(z)\Lambda^2 + a_{2}(z)\id_p)^{-1}$.

\medskip
\noindent{\bf Deriving the bias and variance asymptotics.} Using definition \eqref{mainG}, we can write equation \eqref{eigen2extra} as
\be\label{rewrite X as R} (X_1^\top X_1+X_2^\top X_2)^{-1}=n^{-1}\Sigma_2^{-1/2}V\cal G(0)V^\top\Sigma_2^{-1/2}.\ee
When $z=0$, it is easy to check that equation \eqref{selfomega_a} can be reduced to \eqref{eq_a12extra}, which means that we actually have $a_1(0)=a_1$ and $a_2(0)=a_2$. Hence the matrix limit of $\cal G(0)$ is given by $(a_{1}\Lambda^2 + a_{2}\id_p)^{-1}$. Together with equation \eqref{rewrite X as R} and $V \Lambda^2 V^\top=M^\top M=\Sigma_2^{-1/2}\Sigma_1\Sigma_2^{-1/2}$ by equation \eqref{eigen2}, we can obtain equation \eqref{lem_cov_shift_eq} in the large $n$ limit.

For the bias asymptotics, we can write the left-hand side of equation \eqref{lem_cov_derv_eq} using the derivative of $\cal G$ with respect to $z$ at $z=0$. More precisely, using equation \eqref{eigen2extra} we can obtain that 
\begin{align}
n^2\bignorm{\Sigma_2^{1/2} \bigbrace{ X_1^{\top}X_1 + X_2^{\top}X_2 }^{-1} \Sigma_1^{1/2} w}^2&=n^2 w^\top \Sigma_1^{1/2}\Sigma_2^{-1/2}V\left(   \Lambda U^\top Z_1^\top Z_1 U\Lambda  + V^\top Z_2^\top Z_2V\right)^{-2}V^\top\Sigma_2^{-1/2}\Sigma_1^{1/2}w \nonumber\\
&=  w^\top \Sigma_1^{1/2}\Sigma_2^{-1/2}V\cal G'(0)V^\top\Sigma_2^{-1/2}\Sigma_1^{1/2}w,\label{calculate G'}
\end{align}
where we used equation \eqref{rewrite X as R} in the second step. It is natural to guess that the matrix limit of $\cal G'(0)$ is given by 
\be\label{cal G'0}\cal G'(0) \approx \left.\frac{\dd}{\dd z}\right|_{z=0}(-z\id_p+a_{1}(z)\Lambda^2 + a_{2}(z)\id_p)^{-1} = \frac{\id_p- a_1'(0)\Lambda^2 - a_2'(0)\id_p}{(a_{1}(0)\Lambda^2 + a_{2}(0)\id_p)^2}.\ee
If we let $a_3:=-a_1'(0)$ and $a_4:=-a_2'(0)$, then taking implicit differentiation of equation \eqref{selfomega_a} we can check that $(a_3,a_4)$ satisfies equation \eqref{eq_a34extra}. Then inserting \eqref{cal G'0} into \eqref{calculate G'}, we obtain that 
\begin{align*}
n^2\bignorm{\Sigma_2^{1/2} \bigbrace{ X_1^{\top}X_1 + X_2^{\top}X_2 }^{-1} \Sigma_1^{1/2} w}^2&\approx  w^\top \Sigma_1^{1/2}\Sigma_2^{-1/2}V\frac{a_3\Lambda^2 +(1+ a_4)\id_p}{(a_{1}\Lambda^2 + a_{2}\id_p)^2}V^\top \Sigma_2^{-1/2}\Sigma_1^{1/2}w= w^{\top} \Pi w,
\end{align*}
which concludes \eqref{lem_cov_derv_eq}. Note that in order to have the approximate identity for $\cal G'(0)$ in \eqref{cal G'0}, we not only need to know the asymptotics of $\cal G(0)$, but also need to know the asymptotics of $\cal G(z)$ for general $z$ around $z=0$. This is the main reason why we need to take a general $z$ in the definition of resolvents.

%In the above definition, we have taken the argument of $\cal G$ to be a general complex number, because we will need to use $\cal G'(0)$ in the proof of Lemma \ref{lem_cov_derivative}, which requires a good estimate of $\cal G(z)$ for $z$ around the origin.

%For the variance asymptotic limit, we study the resolvent
%	\[ R(z):= \bigbrace{\Sigma_2^{-1/2}( X_1^{\top}X_1 + X_2^{\top}X_2)\Sigma_2^{-1/2} - z \id}^{-1}, \text{ for any } z\in \C \text{ around } z=0. \]
%Using the techniques from \citet{Anisotropic} and \citet{yang2019spiked}, we find the asymptotic limit of $R(z)$ for any $z$ as $p$ goes to infinity, denoted by $R_\infty(z)$, with an almost optimal convergence rate of $p$.
%In particular, when $z=0$, the asymptotic limit of equation \eqref{lem_cov_shift_eq} is given by
%	\[ \tr[\Sigma_2^{-1/2} \Sigma \Sigma_2^{-1/2}R_\infty(0)]. \]
%
%For the bias asymptotic limit, we show in \todo{where?} that
%$$\bignorm{\Sigma_2^{1/2} (X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} w}^2= w^\top \Sigma_2^{-1/2}R'(0)\Sigma_2^{-1/2} w.$$
%Hence its limit can be calculated through $R_\infty'(z)$, which gives the expression in \eqref{lem_cov_derv_eq}.
%We leave the full proof of Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative} to Appendix \ref{sec_maintools}.
%Combining the above two results, we provide the proof of Theorem \ref{thm_main_informal} in Section \ref{app_proof_main_thm}.

\medskip
\noindent{\bf Gaussian case.} Now we give a heuristic derivation of the matrix limit when the entries of $Z_1$ and $Z_2$ are i.i.d. Gaussian. In this case,
%However, notice that if the entries of $Z_1\equiv Z_1^{\text{Gauss}}$ and $Z_2\equiv Z_2^{\text{Gauss}}$ are i.i.d. Gaussian, then 
by the rotational invariance of the multivariate Gaussian distribution we have
\be\label{eq in Gauss} Z_1 U\Lambda \stackrel{d}{=} Z_1 \Lambda, \quad Z_2 V \stackrel{d}{=} Z_2,\ee
where ``$\stackrel{d}{=}$" means ``equal in distribution". Hence it suffices to consider the following resolvent  
 \begin{equation} \label{resolv Gauss1}
   G(z)= \left( {\begin{array}{*{20}c}
   { -z\id_{p} } & n^{-1/2}\Lambda Z_1^\top & n^{-1/2} Z_2^\top  \\
   {n^{-1/2} Z_1 \Lambda  } & {-\id_{n_1}} & 0 \\
   {n^{-1/2} Z_2} & 0 & {-\id_{n_2}}
   \end{array}} \right)^{-1}.
 \end{equation}

\medskip
\noindent{\bf Schur complements and concentration.} 
%Now we discuss about how to obtain the matrix limit in equation \eqref{defn_piw}.
%\HZ{Divide into several parts like sec 7 to improve readability.}
The core quantities of the derivation are the following resolvent minors, which are defined by removing certain rows and columns of the matrix $H$.

\begin{definition}[Resolvent minors]\label{defn_Minor}
 For any $ (p+n)\times (p+n)$ matrix $\cal A$ and $\fa \in \mathcal I$, we define the minor $\cal A^{(\mathfrak c)}:=(\cal A_{\fa\fb}:\fa,\fb \in \mathcal I\setminus \{\mathfrak a\})$ as the $ (p+n-1)\times (p+n-1)$ matrix obtained by removing the $\mathfrak c$-th row and column in $\cal A$. Note that we keep the names of indices when defining $\cal A^{(\mathfrak c)}$, i.e. $(\cal A^{(\mathfrak c)})_{\fa\fb}= \cal A_{\fa\fb}$ for $\fa,\fb \ne \mathfrak c$. Correspondingly, we define the resolvent minor for $G$ in \eqref{resolv Gauss1} as %(recall equation \eqref{green2})
\begin{align*}
G^{(\mathfrak c)}:&=\left[ \left( {\begin{array}{*{20}c}
   { -z\id_{p} } & n^{-1/2}\Lambda Z_1^\top & n^{-1/2} Z_2^\top  \\
   {n^{-1/2} Z_1 \Lambda  } & {-\id_{n_1}} & 0 \\
   {n^{-1/2} Z_2} & 0 & {-\id_{n_2}}
   \end{array}} \right)^{(\mathfrak c)}\right]^{-1} ,
%= \left( {\begin{array}{*{20}c}
%   { \mathcal G^{(\mathbb T)}} & \mathcal G^{(\mathbb T)} W^{(\mathbb T)}  \\
%   {\left(W^{(\mathbb T)}\right)^\top\mathcal G^{(\mathbb T)}} & { \mathcal G_R^{(\mathbb T)} }  \\
%\end{array}} \right)  ,
\end{align*}
and define the partial traces $m^{(\mathfrak c)}$ and $m^{(\mathfrak c)}_k$, $k=0,1,2,$ by replacing $G$ with $G^{(\mathfrak c)}$ in equation \eqref{defm}. For convenience, we adopt the convention that for the resolvent minor $G^{(\mathfrak c)}$ defined as above, $G^{(\mathfrak c)}_{\fa\fb} = 0$ if $\fa =\mathfrak c$ or $\fb =\mathfrak c$.
\end{definition}

%Note that the resolvent minor $G^{(\mathfrak c)}$ is defined such that it is independent of the entries in the $\mathfrak c$-th row and column of $H$. One will see a crucial use of this fact in the heuristic proof below.

 Using Schur complement formula (see {equation} (\ref{resolvent2})), we have that %for $i \in \cal I_0$, $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\begin{align}
\frac{1}{{G_{ii} }}&=  - z - \frac{\lambda_i^2}{n} \sum_{\mu,\nu\in \mathcal I_1} z_{\mu i}z_{\nu i}G^{\left( i \right)}_{\mu\nu} - \frac{1}{n} \sum_{\mu,\nu\in \mathcal I_2} z_{\mu i}z_{\nu i}G^{\left( i \right)}_{\mu\nu} -2 \frac{\lambda_i}{n} \sum_{\mu\in \cal I_1,\nu\in \mathcal I_2} z_{\mu i}z_{\nu i}G^{\left( i \right)}_{\mu\nu},  \ \ \text{for }\  i \in \cal I_0 , \label{0self_Gii}\\
\frac{1}{{G_{\mu\mu} }}&=  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}\lambda_i \lambda_j z_{\mu i}z_{\mu j} G^{\left(\mu\right)}_{ij}, \quad \frac{1}{{G_{\nu\nu} }}=  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}  z_{\nu i}z_{\nu j}  G^{\left(\nu\right)}_{ij},  \ \ \text{for }\  \mu \in \cal I_1, \ \nu\in \cal I_2, \label{0self_Gmu1} 
\end{align}
where we recall the notations in equation \eqref{labelZ}. For the right-hand side of equation \eqref{0self_Gii}, notice that the resolvent minor $G^{(i)}$ is defined such that it is independent of the entries $z_{\mu i}$ and $z_{\nu i}$. Hence by the concentration inequalities in Lemma \ref{largedeviation}, we have that the  right-hand side of equation \eqref{0self_Gii} concentrates around the partial expectation over the entries $\{z_{\mu i}: \mu \in \cal I_1\cup \cal I_2\}$, i.e., with high probability,
\begin{align*}
\frac{1}{{G_{ii} }}&=  - z - \frac{\lambda_i^2}{n} \sum_{\mu \in \mathcal I_1}  G^{\left( i \right)}_{\mu\mu} - \frac{1}{n} \sum_{\mu\in \mathcal I_2} G^{\left( i \right)}_{\mu\mu} +\oo(1)= - z - \lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1^{(i)}(z)-  \frac{\rho_2}{\rho_1+\rho_2} m_2^{(i)}(z)+\oo(1),  
\end{align*}
where we used the definition of $m_1^{(i)}$ and $m_2^{(i)}$ in equation \eqref{defm} with $G$ replaced by $G^{(i)}$. Intuitively, since we have removed only one column and one row out of the $(p+n)$ columns and rows in $H$, $m_1^{(i)}$ and $m_2^{(i)}$ should be close to the original $m_1$ and $m_2$. Hence we obtain from the above equation that
\begin{align}\label{1self_Gii}
 G_{ii}  = -\left( z +\lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1(z) +  \frac{\rho_2}{\rho_1+\rho_2}m_2(z)+\oo(1)\right)^{-1}.
\end{align}
Similarly, we can obtain from equation \eqref{0self_Gmu1} that for $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\be\label{1self_Gmu} G_{\mu \mu }=-\left(1+\frac{p}{n} m_0 + \oo(1)\right)^{-1},\quad G_{\nu\nu}=-\left(1+\frac{p}{n} m+\oo(1)\right)^{-1},\ee
with high probability.
